### 2020-07-22 paper/introduction.tex
Before submitting to Annals

The \gls{GIF} \cite{doe2002technology} determined six innovative reactor 
systems for further research and commercialization: the Very High-Temperature Reactor (VHTR), the \gls{MSR}, the 
Supercritical Water-Cooled Reactor (SWCR), the Gas-cooled Fast Reactor (GFR), 
the Sodium-cooled Fast Reactor (SFR), and the Lead-cooled Fast Reactor (LFR). 
The MSR is the only liquid-fueled reactor among these reactors. Major nuclear 
centers are pursuing MSRs with renewed interest \cite{betzler_impacts_2019, 
ashraf2020whole}
due to its potential advantages, for example, thorium fuel utilization, small waste production, and ability to use spent fuel.
However, this technology has difficulties including safety, 
online reprocessing, and fuel handling. Safety is one of the primary goals of
the MSR system, and the SD-TMSR, in particular, requires further investigation to support a licensing case.
The unique characteristics of the MSR (liquid fuel, flux level, neutron economy, 
etc.) strongly affect its control system design. Doppler effect and thermal expansion of the fuel reduce
core reactivity when the core heats up, thus MSRs have negative total temperature coefficient of reactivity 
\cite{nuttin2005potential}.
MSR designs have drain tanks to 
hold and cool down the liquid fuel in an emergency. In some MSR design concepts, a freeze plug, which 
is located under the core, passively melts when the fuel temperature reaches a 
critical point, and drains the fuel salt from the reactor vessel to the drain 
tanks. The drain tanks have a subcritical
configuration with a large free 
surface area to readily dissipate heat by passive cooling 
\cite{elsheikh2013safety}.

At the Beginning of Life (BOL), MSRs are commonly loaded with more
fuel than that required to achieve criticality (necessary for long-term core 
operation); this leads to excess reactivity at the BOL. Robertson \emph{et al.} analyzed the \gls{MSBR} loaded by $^{233}$U/$^{232}$Th fuel and concluded that two graphite rods are sufficient for adjusting reactivity during operation due to a small reactivity swing. Robertson \emph{et al.} also reported that 2 safety rods made of boron carbide (B$_4$C) are enough to safely shut the \gls{MSBR} down from any state \cite{robertson_conceptual_1971}. However, the $^{233}$U supply for commercial power plants is not readily available. Our recently published paper \cite{ashraf2020Strategies} concluded that for realistic fueling scenarios (e.g., enriched U/$^{232}$Th, TRU/$^{232}$Th and Pu/$^{232}$Th), large excess reactivity is required. Betzler \emph{et al.} also reported large excess reactivity in Transatomic Power Molten Salt Reactor (TAP MSR) to ensure a 30-year operation cycle \cite{betzler2017assessment}. Additionally, Rykhlevskii \emph{et al.} studied the fuel cycle performance for various Fast-Spectrum \gls{MSR} designs. For a fast MSRs, a considerable excess reactivity is also required for long-term core operation \cite{rykhlevskii_fuel_2019}. Recent studies on MSRs showed that the excess reactivity 
at the BOL may vary from 1.2\% for $^{233}$U-loaded core \cite{rykhlevskii2019modeling,betzler2016modeling} to 8.1\% for Pu/TRU-loaded core \cite{ashraf2020Strategies}.

@@

The SD-TMSR with a thermal power of 2,250 MW$_{th}$ 
\cite{li_optimization_2018} is a graphite-moderated 
molten salt reactor. Adjusting fertile and fissile feed rates helps to control the reactivity of 
the SD-TMSR \cite{ashraf2020Strategies,li_optimization_2018}. However, a reactivity control system for maneuvering and emergency shutdown in the SD-TMSR has not been introduced in the literature \cite{li_optimization_2018,zou2018transition,zhang2020radiotoxicity}. Therefore, the main objective of our study is to introduce a new, 
rapid reactivity control system based on control rods in the SD-TMSR to control 
the reactivity during normal operation and shut down the reactor in case of 
emergency. Six absorbing material options are considered in the context of the 
neutronics and safety parameters in the SD-TMSR. We 
focus on control rod design, absorption ability, integral and differential 
control rod worths, shutdown margin, and shadowing 
effects at steady state calculation.

@@

The online reprocessing and refueling is a distinguishing feature for MSRs comparing with traditional solid fuel reactors. Its neutron characters including $k_{eff}$, excess reactivity, Breeding Ratio (BR), and $^{233}$U production are totally different when operating from those at startup. The $k_{eff}$, excess reactivity, and neutron flux are all related to the scheme of reprocessing. Therefore, the worth of the control rods should be corresponding to the reprocessing scheme; that is, the rate of extraction and addition of elements. However, the present work focuses on the steady-state calculations without taking into account fuel salt depletion.

>>>>

The \gls{GIF} \cite{doe2002technology} determined six innovative reactor 
systems for further research and commercialization: the Very High-Temperature Reactor (VHTR), the \gls{MSR}, the 
Supercritical Water-Cooled Reactor (SWCR), the Gas-cooled Fast Reactor (GFR), 
the Sodium-cooled Fast Reactor (SFR), and the Lead-cooled Fast Reactor (LFR). 
The MSR is the only liquid-fueled reactor among these reactors. Major nuclear 
centers are pursuing MSRs with renewed interest \cite{betzler_impacts_2019, 
ashraf2020whole}
due to its potential advantages, for example, thorium fuel utilization, small waste production, and ability to use spent fuel.
However, this technology has difficulties including safety, 
online reprocessing, and fuel handling. Safety is one of the primary goals of
the MSR system, and the Single-fluid Double-zone Thorium-based Molten Salt Reactor (SD-TMSR), in particular, requires further investigation to support a licensing case.
The unique characteristics of the MSR (liquid fuel, flux level, neutron economy, 
etc.) strongly affect its control system design. Doppler effect and thermal expansion of the fuel reduce
core reactivity when the core heats up, thus MSRs have negative total temperature coefficient of reactivity 
\cite{nuttin2005potential}.
MSR designs have drain tanks to 
hold and cool down the liquid fuel in an emergency. In some MSR design concepts, a freeze plug, which 
is located under the core, passively melts when the fuel temperature reaches a 
critical point, and drains the fuel salt from the reactor vessel to the drain 
tanks. The drain tanks have a subcritical
configuration with a large free 
surface area to readily dissipate heat by passive cooling 
\cite{elsheikh2013safety}.

At the Beginning of Life (BOL), MSRs are commonly loaded with more
fuel than that required to achieve criticality (necessary for long-term core 
operation); this leads to excess reactivity at the BOL. Robertson \emph{et al.} analyzed the \gls{MSBR} loaded by $^{233}$U/$^{232}$Th fuel and concluded that two graphite rods are sufficient for adjusting reactivity during operation due to a small reactivity swing. Robertson \emph{et al.} also reported that two safety rods made of boron carbide (B$_4$C) are enough to safely shut the \gls{MSBR} down from any state \cite{robertson_conceptual_1971}. However, the $^{233}$U supply for commercial power plants is not readily available. Our recently published paper \cite{ashraf2020Strategies} concluded that for realistic fueling scenarios (e.g., enriched U/$^{232}$Th, TRU/$^{232}$Th, and Pu/$^{232}$Th), large excess reactivity is required. Betzler \emph{et al.} also reported large excess reactivity in Transatomic Power Molten Salt Reactor (TAP MSR) to ensure a 30-year operation cycle \cite{betzler2017assessment}. Additionally, Rykhlevskii \emph{et al.} studied the fuel cycle performance for various Fast-Spectrum \gls{MSR} designs. For a fast MSRs, a considerable excess reactivity is also required for long-term core operation \cite{rykhlevskii_fuel_2019}. Recent studies on MSRs showed that the excess reactivity 
at the BOL may vary from 1.2\% for $^{233}$U-loaded core \cite{rykhlevskii2019modeling,betzler2016modeling} to 8.1\% for Pu/TRU-loaded core \cite{ashraf2020Strategies}.

@@

The SD-TMSR with a thermal power of 2,250 MW$_{th}$ 
\cite{li_optimization_2018} is a graphite-moderated 
molten salt reactor. Adjusting fertile and fissile feed rates helps to control the reactivity of 
the SD-TMSR \cite{ashraf2020Strategies,li_optimization_2018}. However, a reactivity control system for maneuvering and emergency shutdown in the SD-TMSR has not been introduced in the literature \cite{li_optimization_2018,zou2018transition,zhang2020radiotoxicity}. Therefore, the main objective of our study is to introduce a new, 
rapid reactivity control system based on control rods in the SD-TMSR to control 
the reactivity during normal operation and shut down the reactor in case of 
emergency. Six absorbing material options are considered in the context of the 
neutronics and safety parameters in the SD-TMSR. We 
focus on control rod design, absorption ability, integral and differential 
control rod worths, shutdown margin, and shadowing 
effects at steady-state calculation.

@@

The online reprocessing and refueling is a distinguishing feature for MSRs comparing with traditional solid fuel reactors. Its neutron characters including effective multiplication factor ($k_{eff}$), excess reactivity, Breeding Ratio (BR), and $^{233}$U production are totally different when operating from those at startup. The $k_{eff}$, excess reactivity, and neutron flux are all related to the scheme of reprocessing. Therefore, the worth of the control rods should be corresponding to the reprocessing scheme; that is, the rate of extraction and addition of elements. However, the present work focuses on the steady-state calculations without taking into account fuel salt depletion.

### 2020-07-22 paper/methods.tex
Before submitting to Annals

The Chinese Academy of Sciences started the program of Thorium Molten Salt Reactor (TMSR) in 2011 \cite{jiang2012advanced}, and many works have been done on the design of MSR and the use of Thorium with the Th-U fuel cycle \cite{li2015analysis}. Among them, the SD-TMSR is proposed in 2018 \cite{li_optimization_2018}, which is a graphite-moderated molten salt reactor with thermal power of 2,250 MW$_{th}$. The design of the SD-TMSR is inspired by the \gls{MSBR} \cite{robertson_conceptual_1971} and the \gls{TMSR} \cite{nuttin2005potential}, which combines the advantages of high breeding ratio (BR) in fast spectrum and low fissile inventory of $^{233}$U in thermal spectrum. SD-TMSR also offers a negative and strong temperature coefficient of reactivity \cite{li_optimization_2018}. Figure~\ref{fig:ff} illustrates the quarter-core view of the
SD-TMSR. The core of the 
SD-TMSR is a right cylinder divided into an inner zone (486 fuel tubes) 
and an outer zone (522 fuel tubes) to enhance breeding performance.
\begin{figure} % replace 't' with 'b' to \centering [hbp!]
	\includegraphics[width=\textwidth]{ff.png}
	\caption{$XZ$ (a) and $XY$ (b) section of the quarter-core model of the 
		SD-TMSR \cite{ashraf2019Preliminary}.}
	\label{fig:ff}
\end{figure}
In this study, the fuel salt composition is LiF-BeF$_2$-(HM)F$_N$ (70-17.5-12.5 mol\%),
where HM is the heavy metal (i.e., $^{232}$Th and fissile 
materials), and $N$ depends on the chosen fissile material and the 
thermochemical state of the liquid fuel salt. Three different types of initial 
fissile materials are considered: (1) $^{233}$U \cite{ashraf2020whole}, 
(2) reactor-grade Pu \cite{marka1993explosive}, and (3) transuranic (TRU) 
elements from \gls{LWR} \gls{SNF} \cite{de2000scenarios}.
The density and volume of the fuel salt are 3.3 g/cm$^{3}$ and 52.9 m$^3$, 
respectively. The liquid fuel salt circulates continuously through the channels
that pierce the graphite hexagonal prisms. The core is surrounded by 
axial and radial graphite reflectors to minimize the neutron leakage.
We adopted the same graphite density reported in the original SD-TMSR paper (2.3 g/cm$^3$) \cite{li_optimization_2018} to be consistent with results in the literature.
A 10-cm-thick B$_4$C cylinder surrounds the core to shield against neutrons and heat.
Finally, the SD-TMSR pressure vessel holds all reactor components and is made of 
a Hastelloy N alloy. The main characteristics of the SD-TMSR are 
summarized in Table~\ref{tab:table1}.

\begin{table}  %[!ht]
	\caption{The main characteristics of the SD-TMSR \cite{li_optimization_2018,ashraf2020whole}.}
	\vspace{0.1in}
	\begin{tabularx}{\textwidth}{l | r}
		\hline
		Thermal power, MW$_{th}$          				&  2,250  \\ 
		Fuel salt components                            & LiF-BeF$_2$-(\gls{HM})F$_N$ \\
		Fuel composition, mole\%                        & 70-17.5-12.5    \\
		$^7$Li enrichment, \%        				& 99.995   \\
		Fuel temperature, K 							& 900  \\
		Dilatation coefficient, g/(cm$^3$$\cdot{}$K)  &  -6.7$\times$10$^{-4}$ \\ 
		Fuel volume, m$^3$  &	52.9 \\
		Fuel density at 900 K, g/cm$^3$		  		& 3.3 \\
		Graphite density, g/cm$^3$             	    & 2.3	\\ 
		Core diameter, cm								& 460  \\
		Core height, cm									& 460  \\
		Side length of the graphite hexagonal prism, cm   & 7.5 \\
		Radius (inner fuel channel), cm							& 3.5  \\
		Radius (outer fuel channel), cm							& 5  \\
		Volume ratio of molten salt to graphite \\in the inner zone	&  0.357  \\
		Volume ratio of molten salt to graphite \\in the outer zone &  1.162  \\

@@

The assessment of the optimal absorber radius should be an important consideration. If the CR radius is too large, the absorber material is not effectively utilized, due to self-shielding. We changed the radius of the CR and calculated the corresponding $k_{eff}$ when all CRs were fully inserted. Figure~\ref{fig:Radius} illustrates the change of the $k_{eff}$ with the radius of the B$_4$C control rod. As shown in Figure~\ref{fig:Radius}, in the region between 0 and 0.75 cm, the $k_{eff}$ decreases sharply with increasing absorber radius. However, in the region between 0.75 and 1.0 cm, there is almost no change in the $k_{eff}$. This is because of the geometry self-shielding of the neutron flux; beyond r = 0.75 cm, the $^{10}$B atoms in the central zone of the CR have relatively low chance for neutron capture.
From the obtained results, the control rod is a cylinder with a radius of 0.75 cm and a height of 520 cm. 
The absorbing material is surrounded by a 0.25-cm-thick cladding made of AIM1 
(15Cr-15Ni) steel alloy \cite{SERAN2017285} and the guide tube is made of 
SiC structural material (see Figure~\ref{fig:cr}). We simulated a 0.1-cm-thick gap between the 
cladding and guide tube to facilitate the control rod movement.
All densities of rod materials are listed in
Table~\ref{tab:table111}.

\begin{table}
	\caption{The densities of the absorbing materials.}
	\vspace{0.05in}
	\begin{center}	
		\begin{tabular}{l  r}  %[!ht]
			%%	\begin{tabularx}{\textwidth}{l | r}
			\hline
			Absorbers & Density [g/cm$^3$]\\
			\hline
			B$_4$C   &  2.54 \\
			HfB$_2$   &  10.5 \\
			HfH$_{1.62}$  &  11.4 \\
			Gd$_2$O$_3$   &  7.04 \\
			Eu$_2$O$_3$  &  7.38 \\
			SiC   & 3.21 \\
			AIM1 cladding   & 7.987 \\
			\hline
			%	\end{tabularx}
		\end{tabular}
		\label{tab:table111}
	\end{center}
\end{table}

@@

Since the total number and distribution of the control assemblies in the SD-TMSR have not been determined, we propose an original distribution as a starting point of this analysis. We added clusters consisting of four control rods to specific graphite hexagonal prisms (elements) in the SD-TMSR core. Every four control rods can move together as one group (cluster). Figure~\ref{fig:graphite_elemen} demonstrates the plan view of the graphite element with the control rods. We propose 25 graphite elements with control rods: 16 Control Safety Devices (CSD) and 9 Shutdown Safety Devices (SSD).

@@

Figure~\ref{fig:core_25} illustrates the numbering scheme of control rod 
clusters in the SD-TMSR core.
The CSD1-16 clusters are represented as yellow and distributed as two rings: inner and outer (peripheral) ring. The inner ring includes CSD from 1 to 6, while the outer ring includes CSD from 7 to 16. Red stands for SSD1-9 clusters.
We distributed the graphite elements with control rod clusters uniformly in 
the inner core of the SD-TMSR, in which the moderator-to-fuel ratio is high.
The selected core segment at the upper left corner of 
Figure~\ref{fig:core_25} shows that both CSD and SSD clusters consist of four 
control rods located at the same distance from the fuel channel center.
\begin{figure}[t!]  % replace 't' with 'b' to \centering
	\centering
	\hspace{+0.65in}
	\includegraphics[width=\textwidth]{core_25.png}
	\caption{Distribution of the graphite elements with CRs in the SD-TMSR core.}
	\label{fig:core_25}
\end{figure}

@@

The effective delayed neutron fraction is calculated by the adjoint-weighted time constants using a perturbation technique in SERPENT-2 \cite{leppanen2014calculation} and it accounts for the steady-state of the fuel salt.

@@

Interference between control rods (CRs) or shadowing effects occur when one 
(or more) control rod impacts the reactivity worth of another control rod in 
the surroundings. 
The insertion of a CR depresses the neutron flux
in its vicinity and makes the curvature of the neutron flux greater. The gradient
of the neutron flux will contrarily increase at a radial distance. The CRs, distributed in the core,
will distort the neutron flux around each other and will impact the reactivity worth.
The term ``shadowing effect" has been used to describe this \cite{oka2014nuclear}.
The shadowing effect appears when the combined rod worth
is less than the sum of the individual worths. Meanwhile, anti-shadowing is observed 
when the combined rod worth is greater than the sum of the individual worths.
The core height-to-diameter ratio (H/D), CR locations, and the three-dimensional configuration 
of the CRs affect the degree of the interference between CRs \cite{oka2014nuclear}.

>>>>

The Chinese Academy of Sciences started the program of Thorium Molten Salt Reactor (TMSR) in 2011 \cite{jiang2012advanced}, and many works have been done on the design of MSR and the use of Thorium with the Th-U fuel cycle \cite{li2015analysis}. Among them, the SD-TMSR is proposed in 2018 \cite{li_optimization_2018}, which is a graphite-moderated molten salt reactor with thermal power of 2,250 MW$_{th}$. The design of the SD-TMSR is inspired by the \gls{MSBR} \cite{robertson_conceptual_1971} and the \gls{TMSR} \cite{nuttin2005potential}, which combines the advantages of high breeding ratio (BR) in fast spectrum and low fissile inventory of $^{233}$U in thermal spectrum. The SD-TMSR also offers a negative and strong temperature coefficient of reactivity \cite{li_optimization_2018}. Figure~\ref{fig:ff} illustrates the quarter-core view of the
SD-TMSR. The core of the 
SD-TMSR is a right cylinder divided into an inner zone (486 fuel tubes) 
and an outer zone (522 fuel tubes) to enhance breeding performance.
\begin{figure} % replace 't' with 'b' to \centering [hbp!]
	\includegraphics[width=\textwidth]{ff.png}
	\caption{$XZ$ (a) and $XY$ (b) section of the quarter-core model of the 
		SD-TMSR \cite{ashraf2019Preliminary}.}
	\label{fig:ff}
\end{figure}
In this study, the fuel salt composition is LiF-BeF$_2$-(HM)F$_N$ (70-17.5-12.5 mol\%),
where HM is the heavy metal (i.e., $^{232}$Th and fissile 
materials), and $N$ depends on the chosen fissile material and the 
thermochemical state of the liquid fuel salt. Three different types of initial 
fissile materials are considered: (1) $^{233}$U \cite{ashraf2020whole}, 
(2) reactor-grade Pu \cite{marka1993explosive}, and (3) transuranic (TRU) 
elements from \gls{LWR} \gls{SNF} \cite{de2000scenarios}.
The density and volume of the fuel salt are 3.3 g/cm$^{3}$ and 52.9 m$^3$, 
respectively. The liquid fuel salt circulates continuously through the channels
that pierce the graphite hexagonal prisms. The core is surrounded by 
axial and radial graphite reflectors to minimize the neutron leakage.
We adopted the same graphite density reported in the original SD-TMSR paper (2.3 g/cm$^3$) \cite{li_optimization_2018} to be consistent with results in the literature.
A 10-cm-thick B$_4$C cylinder surrounds the core to shield against neutrons and heat.
Finally, the SD-TMSR pressure vessel holds all reactor components and is made of 
a Hastelloy N alloy. The main characteristics of the SD-TMSR are 
summarized in Table~\ref{tab:table1}.

\begin{table}  %[!ht]
	\caption{The main characteristics of the SD-TMSR \cite{li_optimization_2018,ashraf2020whole}.}
	\vspace{0.1in}
	\begin{tabularx}{\textwidth}{l | r}
		\hline
		Thermal power, MW$_{th}$          				&  2,250  \\ 
		Fuel salt components                            & LiF-BeF$_2$-(\gls{HM})F$_N$ \\
		Fuel composition, mole\%                        & 70-17.5-12.5    \\
		$^7$Li enrichment, \%        				& 99.995   \\
		Fuel temperature, K 							& 900  \\
		Dilatation coefficient, g/(cm$^3$$\cdot{}$K)  &  -6.7$\times$10$^{-4}$ \\ 
		Fuel volume, m$^3$  &	52.9 \\
		Fuel density at 900 K, g/cm$^3$		  		& 3.3 \\
		Graphite density, g/cm$^3$             	    & 2.3	\\ 
		Core diameter, cm								& 460  \\
		Core height, cm									& 460  \\
		Upper plenum, cm								& 30  \\
		Lower plenum, cm								& 30  \\
		Side length of the graphite hexagonal prism, cm   & 7.5 \\
		Radius (inner fuel channel), cm							& 3.5  \\
		Radius (outer fuel channel), cm							& 5  \\
		Volume ratio of molten salt to graphite \\in the inner zone	&  0.357  \\
		Volume ratio of molten salt to graphite \\in the outer zone &  1.162  \\

@@

The assessment of the optimal absorber radius should be an important consideration. If the CR radius is too large, the absorber material is not effectively utilized, due to self-shielding. We changed the radius of the CR and calculated the corresponding $k_{eff}$ when all CRs were fully inserted. Figure~\ref{fig:Radius} illustrates the change of the $k_{eff}$ with the radius of the B$_4$C control rod. As shown in Figure~\ref{fig:Radius}, in the region between 0 and 0.75 cm, the $k_{eff}$ decreases sharply with increasing absorber radius. However, in the region between 0.75 and 1.0 cm, there is almost no change in the $k_{eff}$. This is because of the geometry self-shielding of the neutron flux; beyond r = 0.75 cm, the $^{10}$B atoms in the central zone of the CR have relatively low chance for neutron capture.
From the obtained results, we proposed the control rod as a cylinder with a radius of 0.75 cm and a height of 520 cm (core height in addition to upper and lower plenums).
The absorbing material is surrounded by a 0.25-cm-thick cladding made of AIM1 
(15Cr-15Ni) steel alloy \cite{SERAN2017285} and the guide tube is made of 
SiC structural material (see Figure~\ref{fig:cr}). We simulated a 0.1-cm-thick gap between the 
cladding and guide tube to facilitate the control rod movement. However, absorber-cladding gap was neglected for simplicity.
All densities of rod materials are listed in
Table~\ref{tab:table111}.

\begin{table}[b!]
	\caption{The densities of the absorbing materials.}
	\vspace{0.05in}
	\begin{center}	
		\begin{tabular}{l  r}  %[!ht]
			%%	\begin{tabularx}{\textwidth}{l | r}
			\hline
			Absorbers & Density [g/cm$^3$]\\
			\hline
			B$_4$C   &  2.54 \\
			HfB$_2$   &  10.5 \\
			HfH$_{1.62}$  &  11.4 \\
			Gd$_2$O$_3$   &  7.04 \\
			Eu$_2$O$_3$  &  7.38 \\
			SiC   & 3.21 \\
			AIM1 cladding   & 7.987 \\
			\hline
			%	\end{tabularx}
		\end{tabular}
		\label{tab:table111}
	\end{center}
\end{table}

@@

Since the total number and distribution of the control assemblies in the SD-TMSR have not been determined, we propose an original distribution as a starting point of this analysis. We added clusters consisting of four control rods to specific graphite hexagonal prisms (elements) in the SD-TMSR core. Every four control rods can move together as one group (cluster). Figure~\ref{fig:graphite_elemen} demonstrates the plan view of the graphite element with the control rods. We propose 25 graphite elements with control rods: 16 Control Safety Devices (CSD) and 9 Shutdown Safety Devices (SSD). We distributed the 25 graphite elements with control rod clusters uniformly in the inner core of the SD-TMSR, in which the moderator-to-fuel ratio is high.

@@

Figure~\ref{fig:core_25} illustrates the numbering scheme of control rod 
clusters in the SD-TMSR core.
The CSD1-16 clusters are represented as yellow and distributed as two rings: inner and outer (peripheral) ring. The inner ring includes CSD from 1 to 6, while the outer ring includes CSD from 7 to 16. Red stands for SSD1-9 clusters.
The selected core segment at the upper left corner of 
Figure~\ref{fig:core_25} shows that both CSD and SSD clusters consist of four 
control rods located at the same distance from the fuel channel center.
\begin{figure}[t!]  % replace 't' with 'b' to \centering
	\centering
	\hspace{+0.65in}
	\includegraphics[width=\textwidth]{core_25.png}
	\caption{Distribution of the graphite elements with CRs in the SD-TMSR core.}
	\label{fig:core_25}
\end{figure}

@@

The effective delayed neutron fraction is calculated by the adjoint-weighted time constants using a perturbation technique in SERPENT-2 \cite{leppanen2014calculation} and it accounts for the steady state of the fuel salt.

@@

Interference between control rods (CRs) or shadowing effects occur when one 
(or more) control rod impacts the reactivity worth of another control rod in 
the surroundings. 
The insertion of a CR depresses the neutron flux
in its vicinity and makes the curvature of the neutron flux greater. The gradient
of the neutron flux will contrarily increase at a radial distance. The CRs, distributed in the core,
will distort the neutron flux around each other and will impact the reactivity worth.
The term ``shadowing effect" has been used to describe this phenomenon \cite{oka2014nuclear}.
The shadowing effect appears when the combined rod worth
is less than the sum of the individual worths. Meanwhile, anti-shadowing is observed 
when the combined rod worth is greater than the sum of the individual worths.
The core height-to-diameter ratio (H/D), CR locations, and the three-dimensional configuration 
of the CRs affect the degree of the interference between CRs \cite{oka2014nuclear}.

### 2020-07-22 paper/results.tex
Before submitting to Annals

The SERPENT-2 is used to obtain the presented results.
We use the ENDF-VII.0 cross section library for all calculations in the current work. The statistical error in $k_{eff}$ is $\leq$ $25$ $pcm$.
The initial calculation state of a full 3D model of the SD-TMSR is identified by normal operation 
conditions (see Table~\ref{tab:table1}) and fully withdrawn control rod 
clusters. 
To validate the proposed control rods 
system we adopted the same operation conditions (as in the initial calculation 
state) and changed the position of the control rod clusters along the $Z$ 
direction.
The obtained results including reactivity, control rod 
worth, and interference effects (shadowing effects) are discussed below.

@@

The excess reactivity $\rho$$_e$ is calculated at zero burnup (steady-state 
calculation), when all control rods (CRs) are fully withdrawn. The $\rho_e$ for $^{233}$U, 
reactor-grade Pu, and TRU used as initial fissile material are listed in Table~\ref{tab:excess}.
For $^{233}$U case \cite{ashraf2020Strategies} the maximum observed excess reactivity is $4.27\pm0.01$ $\$$ during 60 effective full-power years (EFPY) of operation (see Figure~\ref{fig:keff_25} at $\approx$ 8 EFPY).
The proposed reactivity control system must compensate such reactivity at startup and during burnup.
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{keff_25.png}
	\vspace{-0.5in}
	\caption{Uncontrolled effective multiplication factor during 60 EFPY of reactor operation
		including periodic $^{233}$U/$^{232}$Th insertion (confidence interval $\pm\sigma$ is shaded) \cite{ashraf2020whole}.} 
	\label{fig:keff_25}
\end{figure}

@@

@@

The worth of the Control Safety Devices (CSD) clusters is $1.56$ times greater than 
the worth of the Shutdown Safety Devices (SSD) system for all absorbing materials. Either CSD or SSD 
clusters are able to shut down the reactor initially loaded by 
$^{233}$U and reactor-grade Pu regardless the absorbing material type.
However, only SSD clusters made of B$_4$C-90 are able to shut down the SD-TMSR 
initially loaded by transuranic (TRU) elements ($\rho_e$ is $15.38\pm0.04$ $\$$).
The reason for this is the much larger 
absorption cross section of $^{10}$B in the relatively soft neutron energy 
spectrum of the SD-TMSR core started with TRU \cite{ashraf2020Strategies}.

@@

We separately calculated the worth of CSD2, CSD9, SSD1, and SSD4 clusters (i.e., clusters located in the center and at the boundary between the core
zones, see Figure~\ref{fig:core_25}) to investigate the variation of CRW with the position in the inner core.
The CRW decreases in the direction of the outer core zone. The outer core zone 
has smaller moderator-to-fuel ratio ($0.86$) compared with the central zone 
($2.8$); consequently, the neutron energy spectrum is faster in the peripheral 
zone than in the center of the core degrading the CR's 
absorption ability in that region.

@@

The Shutdown Safety Devices (SSD) clusters are designed mainly for an emergency shutdown, thus it should 
provide the reactor core with sufficient negative reactivity. The 
shutdown margin (SDM) is calculated by equation~\ref{Equ:6}.
Table~\ref{tab:table2} summarizes the shutdown margins for the SD-TMSR core 
initially loaded with $^{233}$U, reactor-grade Pu, and TRU 
elements for different absorbing materials. All absorbing materials provide a sufficient positive
SDM for the SD-TMSR core that is initially loaded with 
$^{233}$U and reactor-grade Pu. However, the SDMs for the TRU case are 
negative or slightly positive (in B$_4$C-90 case); this makes the SSD clusters 
unreliable to shut down the reactor loaded with TRU.
As mentioned previously, the SSD clusters should 
provide a sufficient positive SDM.

@@

The amplification factor (A$_{CRi}$) results show that the CSD, SSD, CSD inner 
ring, and SSD1 are slightly amplified due to the anti-shadowing effects. The 
anti-shadowing is observed when the combined rod worth is greater than the sum 
of the individual worths. As listed in Tables~\ref{tab:worth} and ~\ref{tab:table25}, the strongest anti-shadowing effect occurred in 
SSD4 and CSD9 clusters that are located at the boundary between the core zones 
with different moderator-to-fuel ratios. This can be attributed to the fact that the SSD4 and CSD9 clusters are surrounded by fewer clusters compared with other clusters located in the inner zone of the SD-TMSR (see Figure~\ref{fig:core_25}). Consequently, low interference between the SSD4 and CSD9 clusters and other surrounding clusters is observed. The obtained results show a negligible relationship between the absorbing material type and the interference between the CR clusters (i.e., the A$_{CRi}$).

>>>>

The SERPENT-2 is used to obtain all presented results.
We use the ENDF-VII.0 cross section library for all calculations in the current work. The statistical error in $k_{eff}$ is $\leq$ $25$ $pcm$.
The initial calculation state of a full 3D model of the SD-TMSR is identified by normal operation 
conditions (see Table~\ref{tab:table1}) and fully withdrawn control rod 
clusters. 
The obtained results including excess reactivity, control rod 
worth, shutdown margin, interference effects (shadowing effects), etc. are discussed below.

@@

The excess reactivity $\rho$$_e$ is calculated by using equation~\ref{Equ:1} at zero burnup (steady-state 
calculation), when all control rods (CRs) are fully withdrawn. The $\rho_e$ for $^{233}$U, 
reactor-grade Pu, and TRU used as initial fissile material are listed in Table~\ref{tab:excess}.
For $^{233}$U case \cite{ashraf2020Strategies} the maximum observed excess reactivity is $4.27\pm0.01$ $\$$ during 60 effective full-power years (EFPY) of operation (see Figure~\ref{fig:keff_25} (at $\approx$ 8 EFPY)).
The proposed reactivity control system must compensate such reactivity at startup and during burnup.

@@

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{keff_25.png}
	\vspace{-0.5in}
	\caption{Uncontrolled effective multiplication factor during 60 EFPY of reactor operation
		including periodic $^{233}$U/$^{232}$Th insertion (confidence interval $\pm\sigma$ is shaded) \cite{ashraf2020whole}.} 
	\label{fig:keff_25}
\end{figure}

@@

The worth of the Control Safety Devices (CSD) clusters is $1.56$ times greater than 
the worth of the Shutdown Safety Devices (SSD) system for all absorbing materials. Either CSD or SSD 
clusters are able to shut down the reactor initially loaded by 
$^{233}$U and reactor-grade Pu regardless of the absorbing material type.
However, only SSD clusters made of B$_4$C-90 are able to shut down the SD-TMSR 
initially loaded by transuranic (TRU) elements ($\rho_e$ is $15.38\pm0.04$ $\$$).
The reason for this is the much larger 
absorption cross section of $^{10}$B in the relatively soft neutron energy 
spectrum of the SD-TMSR core started with TRU \cite{ashraf2020Strategies}.

@@

We separately calculated the worth of CSD2, CSD9, SSD1, and SSD4 clusters (i.e., clusters located in the center and at the boundary between the core
zones, see Figure~\ref{fig:core_25}) to investigate the variation of CRW with the position in the inner core.
The CRW decreases in the direction of the outer core zone. The outer core zone 
has smaller moderator-to-fuel ratio ($0.86$) compared with the central zone 
($2.80$); consequently, the neutron energy spectrum is faster in the peripheral 
zone than in the center of the core degrading the CR's 
absorption ability in that region.

@@

The Shutdown Safety Devices (SSD) clusters are designed mainly for an emergency shutdown, thus it should 
provide the reactor core with sufficient negative reactivity. The 
shutdown margin (SDM) is calculated by equation~\ref{Equ:6}.
Table~\ref{tab:table2} summarizes the shutdown margins for the SD-TMSR core 
initially loaded with $^{233}$U, reactor-grade Pu, and TRU 
elements for different absorbing materials. All absorbing materials provide a sufficient positive
SDM for the SD-TMSR core that is initially loaded with 
$^{233}$U and reactor-grade Pu. However, the SDMs for the TRU case are 
negative or slightly positive (in B$_4$C-90 case); this makes the SSD clusters 
unreliable to shut down the reactor loaded with TRU.
For the TRU case, increasing the number of SSD clusters would provide a sufficient positive SDM.
This should be carefully studied when designing the reactivity control system.

@@

The amplification factor (A$_{CRi}$) results show that the CSD, SSD, CSD inner 
ring, and SSD1 are slightly amplified due to the anti-shadowing effects. The 
anti-shadowing is observed when the combined rod worth is greater than the sum 
of the individual worths. As listed in Tables~\ref{tab:worth} and ~\ref{tab:table25}, the strongest anti-shadowing effect occurred in 
SSD4 and CSD9 clusters that are located at the boundary between the core zones 
with different moderator-to-fuel ratios. This can be attributed to the fact that the SSD4 and CSD9 clusters are surrounded by fewer clusters compared with other clusters located in the inner zone of the SD-TMSR (see Figure~\ref{fig:core_25}). Consequently, low interference between these SSD4 and CSD9 clusters and other surrounding clusters is observed. The obtained results show a negligible relationship between the absorbing material type and the interference between the CR clusters (i.e., the A$_{CRi}$).
